---
title: "Handon_Ex09"
editor: visual
---

# 

Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method

## Overview

Geographically weighted regression (GWR) is a spatial statistical technique that takes non-stationary 非稳态的 variables into consideration ( climate, demographic factors) In this hand-on, I will build [hedonic](https://www.investopedia.com/terms/h/hedonicpricing.asp) pricing model by using GWR methods

## The Data

Two data sets will be used in this model building exercise, they are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)

-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

## Getting Started

### Load necessary Package

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

### Importing geospatial data

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

### Updating CRS information

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

```{r}
st_crs(mpsz_svy21)
```

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

### 
